---
title: "WUP_04"
author: "Cleo Ware"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata<- read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using ChatGPT and CandP helper files to compare numerical data and catagorical data in the GSS data set. We will compare WORDSUM and how it effected by Education level. 


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE,echo=FALSE}
ggplot(data = mydata, aes(x = DEGREE, y = WORDSUM) ) +
geom_boxplot ( notch=FALSE , outlier.colour = "red" , outlier.size = 2 , fill = "darkgray" ) +
stat_boxplot ( geom = 'errorbar' ) +
labs ( x = "DEGREE" , y = "WORDSUM" , title = "WORDSUM EXPLAINED BY DEGREE" ) +
theme ( plot.title = element_text (hjust = 0.5, size = 16) )
```

It appears that the entire IQR of the HIgh School is underneath the median of all of the college groups.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE,echo=FALSE}
favstats( WORDSUM ~ DEGREE , data= mydata )
```

The mean and median of the LT high school and the high school subcategories are far lower than the mean and medians of the university level data points.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r include=TRUE,echo=FALSE}
model1 <- aov( WORDSUM ~ DEGREE , data = mydata )
conf_interval <- confint(model1, level = 0.98)
summary.aov(model1)
```

Ran a confidence interval test and the above results show a much higher mean for DEGREE.

```{r include=TRUE,echo=FALSE}
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1,conf.level = 0.98),las=2)
```

I ran TukeyHSD test 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The analysis of the 2018 General Social Survey (GSS) data provides strong evidence to reject the null hypothesis ($H_0$) that there is no difference between the average score on the WORDSUM test and the level of schooling (DEGREE). Instead, the data suggests a significant difference in average WORDSUM scores between at least two levels of DEGREE.

Specifically, the mean and median WORDSUM scores of individuals with lower education levels, such as "LT HIGH SCHOOL" and "HIGH SCHOOL," are notably lower than those of individuals with higher education levels, such as "BACHELOR" and "GRADUATE." This implies that education level has a significant impact on WORDSUM scores, and individuals with higher education levels tend to have higher scores on the WORDSUM test.

These findings underscore the importance of education in cognitive abilities and have implications for policy and educational interventions aimed at improving cognitive skills among individuals with lower levels of education. Further research may explore the underlying factors contributing to these differences in WORDSUM scores among education levels.

